### 2018-04-07 Manuscript/chapter4.tex
Images

Although there is growing interest in how viruses and diet quality affect the health and sustainability of honeybees, as well as a recognition that such factors might operate interactively, there are only a small number of experimental studies thus far directed toward elucidating the interactive effects of these two factors in honeybees (\citealt{intNV}, \citealt{intNV2}, \citealt{intNV3}). We recently used laboratory cages and nucleus hive experiments to investigate how these two factors interact (\citealt{adamInt}). As part of our experiment, we specifically studied the interactive effects of IAPV infection and monofloral diet quality on titers on honeybee mortality rates.

There are several reasons why, in part of our study, we focused only on diet quality (monofloral diets) as opposed to diet diversity (monofloral diets versus polyfloral diets). First, when assessing diet diversity, a sugar diet is often used as a control. However, such an experimental design does not reflect real-world conditions for honeybees as they rarely face a total lack of pollen (\citealt{DiPasquale}). Second, in studies that compared honeybee health using monofloral and polyfloral diets at the same time, if the polyfloral diet and one of the high-quality monofloral diets both exhibited similarly beneficial effects, then it was difficult for the authors to assess if the polyfloral diet was better than most of the monofloral diets because of its diversity or because it contained as a subset the high-quality monofloral diet (\citealt{DiPasquale}). Third, colonies used for pollination in agricultural areas (monoculture) face less diversified pollens (according to Brodschneider, 2010). Pollinating areas are currently undergoing landscape alteration and agriculture intensification, and bees are increasingly faced wtih less diversified diets (monoculture) (\citealt{landscape1}, \citealt{brodschneider}). As a result, there is a need to better understand how monofloral diets affect honeybee health as a step toward mitigating the negative impact of human activity on the honeybee population.

@@

>>>>

Although there is growing interest in how viruses and diet quality affect the health and sustainability of honeybees, as well as a recognition that such factors might operate interactively, there are only a small number of experimental studies thus far directed toward elucidating the interactive effects of these two factors in honeybees (\citealt{intNV}, \citealt{intNV2}, \citealt{intNV3}). We recently used laboratory cages and nucleus hive experiments to investigate how these two factors interact (\citealt{adamInt}). As part of our experiment, we specifically studied the interactive effects of IAPV infection and monofloral diet quality on titers and honeybee mortality rates.

There are several reasons why, in this part of our previous study, we focused only on diet quality (monofloral diets) as opposed to diet diversity (monofloral diets versus polyfloral diets). First, when assessing diet diversity, a sugar diet is often used as a control. However, such an experimental design does not reflect real-world conditions for honeybees as they rarely face a total lack of pollen (\citealt{DiPasquale}). Second, in studies that compared honeybee health using monofloral and polyfloral diets at the same time, if the polyfloral diet and one of the high-quality monofloral diets both exhibited similarly beneficial effects, then it was difficult for the authors to assess if the polyfloral diet was better than most of the monofloral diets because of its diversity or because it contained as a subset the high-quality monofloral diet (\citealt{DiPasquale}). Third, colonies used for pollination in agricultural areas (monoculture) face less diversified pollens (according to Brodschneider, 2010). Pollinating areas are currently undergoing landscape alteration and agriculture intensification, and bees are increasingly faced wtih less diversified diets (monoculture) (\citealt{landscape1}, \citealt{brodschneider}). As a result, there is a need to better understand how monofloral diets affect honeybee health as a step toward mitigating the negative impact of human activity on the honeybee population.

Consequently, in our prior study, for our nutrition factor, we examined two monofloral pollen diets, Cistus (Rockrose) and Castanea (Chestnut). Cistus pollen is generally considered less nutritious than Castanea pollen due to its lower levels of protein, amino acids, antioxidants, calcium, and iron (\citealt{DiPasquale}, \citealt{adamInt}). For our virus factor, one level contained bees that were infected with IAPV and another level contained bees that were not infected with IAPV. This experimental design resulted in four treatment groups that allowed us to assess main effects and interactive effects between diet quality and IAPV infection in honeybees.

We discovered that the higher-quality Castanea diet had the ability to significantly reduce mortality in the presence of IAPV infection (Figure \ref{fig:mortalityAll}).

Mortality of bees 72 h post-inoculation differed among the treatment groups (mixed model ANOVA across all treatment groups, df=7, 108; F=19.28; p<0.0001). There was a significant effect of both virus treatment (mixed model ANOVA, df=1, 108; F=95.26; p<0.0001) and diet treatment (mixed model ANOVA, df=3,108; F=9.93; p<0.001) with a significant interaction between the two factors (mixed model ANOVA, df=3, 108; F=3.311, p=0.023). The virus treatment was effective-- in all cases, virus-treated bees showed significantly higher mortality than bees from cages fed the same diet but no virus (Tukey HSD, p<0.05). Without virus exposure, none of the treatment groups showed differences in mortality (Tukey HSD, p>0.05). However, between the virus-treated groups, both bees fed Castanea or polyfloral pollen showed significantly lower mortality than bees fed no pollen (Tukey HSD, p<0.05); bees fed Cistus pollen exhibited intermediate mortality (Tukey HSD, p<0.05; Figure 1). Overall, we find that virus infection increases mortality, but some pollen diets are capable of significantly reducing the severity of this effect.

glht (General Linear Hypotheses)

\begin{figure}[H]
\centering
  \begin{framed}
  \includegraphics[width=0.8\textwidth]{Images/mortalityAll}
  \end{framed}
  \caption{Mortality rates between the four treatments. Each treatment contains 15 samples (cages).}
  \label{fig:mortalityAll}
\end{figure}

\begin{figure}[H]
\centering
  \begin{framed}
  \includegraphics[width=0.8\textwidth]{Images/mortalityRNASeq}
  \end{framed}
  \caption{Mortality rates between the four treatments. Out of the original 15 samples (cages), each treatment here only contains the 6 samples (cages) that were used for RNA-sequencing.}
  \label{fig:mortalityRNASeq}
\end{figure}

\begin{figure}[H]
\centering
  \begin{framed}
  \includegraphics[width=0.8\textwidth]{Images/IAPVAll}
  \end{framed}
  \caption{IAPV titers between the four treatments. Each treatment contains 15 samples (cages).}
  \label{fig:IAPVAll}
\end{figure}

\begin{figure}[H]
\centering
  \begin{framed}
  \includegraphics[width=0.8\textwidth]{Images/IAPVRNASeq}
  \end{framed}
  \caption{IAPV titers between the four treatments. Out of the original 15 samples (cages), each treatment here only contains the 6 samples (cages) that were used for RNA-sequencing.}
  \label{fig:IAPVRNASeq}
\end{figure}

Summary of what that paper found mortality, micronutrients
 We found that high quality diets (polyfloral pollen and high quality single-source pollen) have the potential to reduce mortality in the face of infection with Israeli acute paralysis virus (IAPV)
- There was a significant interaction between diet and virus infection on mortality, with associated differences in bee virus titers, suggesting good diets can help bees keep viral infection levels down

@@

RNA-seq
RNA prep
Illumina sequencing at DNA facility
What version of genome mapped to
What mapping program did he use
What percent of reads mapped to the genome
Statistics (DESeq, EdgerR, visualizations, overlaps fisher, GO)
